Prediction of Periprosthetic Joint Infection after Total Hip Arthroplasty: a simplified and externally validated model based on observational registry data from Sweden and ...

### 2020-06-04 reports/article.Rmd
version sent to NH and ABP

get_auc <- function(Time, model, version = "1") {
  data_auc_ci %>% 
  filter(time == Time, Model == model) %>% 
  select(ends_with(version)) %>% 
  pluck(1)
}
```

@@

Model development was performed by bootstrap ranking and a logistic least absolute shrinkage and selection operator (LASSO).[@Guo2015; @Baranowski2020] Age was normalized prior to modeling, and the estimated coefficients thus indicated variable importance on the same scale as categorical variables. In order to account for the potentially U-shaped risk of developing PJI associated with either low or high BMI, this variable was treated as a categorical variable divided into four BMI intervals based on the six intervals propsed by the World Health Organization (WHO; $<25$: under- or normal weight, $25-29.9$: overweight, $30-34.9$: class I obesity, $\geq 35$: class II-III obesity).

For each outcome (PJI within 90 days or 2 years), 1,000 bootstrap samples were drawn from the observed data set.[@Austin2004] Ten-fold cross validation was performed for every bootstrap sample with a range of potential penalty values ($\lambda$:s) in a logistic LASSO model, keeping only $\lambda$:s that minimized the mean cross-validated deviances in each sample and using these $\lambda$:s to estimate model coefficients for each potential predictor. Those coefficients were ranked based on their absolute values as a measure of variable importance. Piece-wise linear regression was applied to find a break-point, and influential predictors with absolute coefficients value above this pointm, scored a point. The whole process was repeated 100 times. Potential predictors that were selected at least once were used in main effects multivariable logistic regression models without penalty and without pre-normalization of numeric variables. Even more parsimonious models with potential predictors chosen at least 33 out of 100 times were evaluated for comparison. Univariable prediction models based on either ASA class, the Charlson, Elixhauser, or the Rx Risk V co-morbidity indices were fitted for comparison. Odds ratios for the final models were estimated with 95 % confidence intervals.

@@

Patients with `r excl_factors` were uncommon, wherefore those co-morbidities were excluded as potentisl predictors prioir to further modeling. ROC-curves (Fig. \@ref(fig:rocs)) and AUC-values (Fig. \@ref(fig:aucci)) for the main and reduced models were no different. For simplicity, we therefore focus on the reduced models. Estimated probability density curves for patients with and without PJI were partially overlapping, although patients with PJI had on average higher predicted probabilites for such event (Fig. \@ref(fig:sep)).

@@

```{r tab1}
source("../lib/clean_names.R")

>>>>

get_auc <- function(Time, model, version = "1") {
  data_auc_ci %>% 
  filter(Period == Time, Model == model) %>% 
  select(ends_with(version)) %>% 
  pluck(1)
}
```

@@

Model development was performed by bootstrap ranking and a logistic least absolute shrinkage and selection operator (LASSO).[@Guo2015; @Baranowski2020] Age was normalized prior to modeling, and the estimated coefficients thus indicated variable importance on the same scale as categorical variables. BMI divided into four intervals based on the six intervals proposed by the World Health Organization (WHO; $<25$: under- or normal weight, $25-29.9$: overweight, $30-34.9$: class I obesity, $\geq 35$: class II-III obesity).

For each outcome (PJI within 90 days or 2 years), ten-fold cross validation was performed with a range of potential penalty values ($\lambda_i$:s) in a logistic LASSO regression model. We kept the $\lambda_i$ that minimized the mean cross-validated deviance. 100 bootstrap samples were then drawn from the observed data set.[@Austin2004]

Logistic LASSO regression was performed for each sample using the chosen $\lambda_i$ as penalty term. Regression coefficients were ranked based on their absolute values as a measure of variable importance. We then used Bayesian methods to estimate a posterior distribution and to identify two change points in a linear segmented piecewise regression model. The second change point was then used as a break-point. Influential predictors with absolute coefficients value above this point, scored a point. The whole process, each with 100 Bootstrap replicates, was repeated 100 times (yielding 10,000 bootstrap samples in total). Potential predictors that were selected at least once out of 100 were then used in main effects multivariable logistic regression models without penalty and without pre-normalization of numeric variables. Even more parsimonious models with potential predictors chosen at least 50 out of 100 times were evaluated for comparison. Univariable prediction models based on either ASA class, the Charlson, Elixhauser, or the Rx Risk V co-morbidity indices were fitted for comparison. Odds ratios for the final models were estimated with 95 % confidence intervals.

@@

Patients with `r excl_factors` were uncommon, wherefore those co-morbidities were excluded as potential predictors prior to further modeling. ROC-curves (Fig. \@ref(fig:rocs)) and AUC-values (Fig. \@ref(fig:aucci)) for the main and reduced models were no different. For simplicity, we therefore focus on the reduced models. Estimated probability density curves for patients with and without PJI were partially overlapping, although patients with PJI had on average higher predicted probabilities for such event (Fig. \@ref(fig:sep)).

@@

```{r tab1, warning=FALSE}
source("../lib/clean_names.R")
